Powering Hidden Markov Model by Neural Network based Generative Models

### 2019-09-01 manuscript/main.tex
add noise testing

The problem \eqref{eq:em-m-opt} can be reformulated as
\begin{align}\label{eq:m-step-subs}
  &\umax{\bm{H}} \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) \nonumber \\
  =&\umax{\bm{q}}\Qq(\bm{q}; \bm{H}^{\mathrm{old}}) + \umax{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}}) 
     + \umax{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}),
\end{align}
where
\begin{align}
  \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) 
  % &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p({s}_{1})  \right] \nonumber\\
  &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p({s}_{1};\bm{H})  \right], \label{eq:init-distribution-update}\\
  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \sum_{t=1}^{T}\log\,p({s}_{t+1}|{s}_{t}; \bm{H}) \right], \label{eq:transition-update}\\
  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p(\ubm{x}, \ubm{\kappa}| \ubm{s}; \bm{H}) \right]. \label{eq:generative-model-update}
\end{align}

@@

For a flow model, let us assume that
the feature $\bm{h}_l$ at the $l$'th layer has two subparts as
$\bm{h}_l = [\bm{h}_{l,a}^{\intercal} \, , \, \bm{h}_{l,b}^{\intercal}]^{\intercal}$. The efficient invertible mapping of flow model comes from following forward and inverse relations between $(l-1)$'th and $l$'th layers
\begin{align}\label{eq-gl}
    \bm{h}_{l} &=
    \begin{bmatrix}
      \bm{h}_{l,a}\\
      \bm{h}_{l,b}
    \end{bmatrix}
    =
    \begin{bmatrix}
      \bm{h}_{l-1,a}\\
      \left(  \bm{h}_{l-1,b} - \bm{m}_b(\bm{h}_{l-1,a}) \right)\oslash \bm{m}_a(\bm{h}_{l-1,a}) 
    \end{bmatrix}, \nonumber\\
        \bm{h}_{l-1} \hspace{-0.1cm}&=\hspace{-0.1cm}
    \begin{bmatrix}
      \bm{h}_{l-1,a}\\
      \bm{h}_{l-1,b}
    \end{bmatrix}
    \hspace{-0.1cm}=\hspace{-0.1cm}
    \begin{bmatrix}
      \bm{h}_{l,a}\\
      \bm{m}_a(\bm{h}_{l,a})\odot \bm{h}_{l,b} + \bm{m}_b(\bm{h}_{l,a})
    \end{bmatrix},
\end{align}
where $\odot$ denotes element-wise product, $\oslash$ denotes
element-wise division, and $\bm{m}_a(\cdot), \bm{m}_b(\cdot)$ can be
complex non-linear mappings (implemented by neural networks).
For the flow-based neural network, the determinant of Jacobian matrix is
\begin{equation}\label{eq:cat-jacobian}
  \begin{array}{rl}
    \mathrm{det}(\nabla{\bm{f}_{s,\kappa}}) & = \prod_{l=1}^L \det (\nabla{\bm{f}_{s,\kappa}^{[l]}}),
  \end{array}
\end{equation}
where $\nabla{f_{s,\kappa}^{[l]}}$ is the Jacobian of the transformation from the $l$-th layer to the $(l-1)$-th layer, i.e., the inverse transformation. We compute the determinant of the Jacobian matrix as
\begin{align}\label{eq-hl-determinate}
  \det (\nabla{f_{s,\kappa}^{[l]}})& = \det \left[  \pd{\bm{h}_{l-1}}{\bm{h}_l} \right] \nonumber\\
                                   & = \det
                                     \begin{bmatrix}
                                       \bm{I}_a & \mathbf{0} \nonumber\\
                                       \pd{\bm{h}_{l-1,b}}{\bm{h}_{l,a}} & \mathrm{diag}(\bm{m}_a(\bm{h}_{l,a}))
                                     \end{bmatrix}\nonumber\\
                                   &= \det \left( \mathrm{diag}(\bm{m}_a(\bm{h}_{l,a})) \right),
\end{align}
where $\bm{I}_a$ is identity matrix and $\mathrm{diag}(\cdot)$ returns a square matrix with the elements of $(\cdot)$ on the main diagnal.
% \textcolor{red}{do I still need this?}
% Then the pdf is
% \begin{align}
%   p(\bm{x}) & =  p(\bm{z}) \big| \mathrm{det}(\bm{J}) |_{\bm{z}={\bm{f}}({\bm{x}})}\big| \nonumber\\
%             &  = p(\bm{z}) \prod_{l=1}^L \abs{\det\left( \mathrm{diag}(\bm{m}_a(\bm{h}_{l,a}))]  \right)}.
                %   \end{align}

@@

$p(\smtr{s}{1}{r};\bm{H})$ is the probability of initial state of GenHMM for $r$-th sequential sample. Actually $q_i = p({s}_{1} =i;\bm{H}) $, $i= 1, 2, \cdots, |\Ss|$. Solution to the problem
\begin{align}
  \bm{q} &= \uargmax{\bm{q}} \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) \nonumber \\
         &\mathrm{s.t.} \sum_{i=1}^{ |\Ss| }q_i = 1, q_i \geq 0, \forall i.
\end{align}
is
\begin{equation}\label{eq:update-initial-state-prob}
  q_i = \frac{1}{R} \sum_{r=1}^{R} p(\smtr{s}{1}{r}=i | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}), \forall\; i = 1, 2, \cdots, |\Ss|.
\end{equation}

@@

Since $\bm{A}_{i, j}  = p(\smtr{s}{t+1}{r}=j|\smtr{s}{t}{r}=i; \bm{H})$ is the element of transition matrix $\bm{A}$, the solution to the problem
\begin{align}\label{eq:update-transition-prob}
  \bm{A} = &\uargmax{\bm{A}} \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) \nonumber \\
  \mathrm{s.t.} &\hspace{0.2cm} \bm{A} \cdot \bm{1} = \bm{1} \nonumber \\
           & \bm{A}_{i,j} \geq 0 \,\, \forall i,j,
\end{align}
is
\begin{equation}
  \bm{A}_{i,j} = \frac{\bar{\xi}_{i,j}}{\sum_{k = 1}^{|\Ss|} \bar{\xi}_{i,k}},
\end{equation}
where
\begin{equation}\label{eq:update-transition-solt}
  \bar{\xi}_{i,j} = \sum_{r= 1}^{R} \sum_{t= 1}^{T^{r}-1}{p(\smtr{s}{t}{r}=i, \smtr{s}{t+1}{r}=j| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}.
\end{equation}
\subsection{On Convergence of GenHMM}
In pursuit of representing a dataset by GenHMM,  we are interested if the learning solution discussed in subsection~\ref{subsec:optmGenHMM} would convergence. The properties on GenHMM's convergence is analyzed as follows.

@@

\subsection{Algorithm of GenHMM}

@@

\begin{algorithm}[H]
  \caption{Learning of GenHMM}\label{algo:genhmm}
  \begin{algorithmic}[1]
    \STATE {\bfseries Input:}{
      Initializing $\bm{H}^{\mathrm{old}}, \bm{H} \in \Hh$ gives: \\
      $\bm{H}^{\mathrm{old}} = \{\Ss, \bm{q}^{\mathrm{old}}, A^{\mathrm{old}}, p(\bm{x}|s; \bm{\Phi}^{\mathrm{old}})\}$, $\bm{H} = \{\Ss, \bm{q}, A, p(\bm{x}|s; \bm{\Phi})\}$,\\
      empirical distribution $\hat{p}(\bm{x})$ of dataset;} \\
    \STATE Initialize $\bm{H}$
    \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
    \STATE Set learning rate $\eta$, neural network optimization epochs $N$ per EM step
    \FOR { $\bm{H}$ not converge}
    \FOR {epoch $n < N$}
    \STATE Sample a batch of data $\left\{ \ubmr{x}{r} \right\}_{r=1}^{R_b}$ from dataset $\hat{p}(\ubm{x})$ with batch size $R_b$

@@

The TIMIT dataset consists of $5300$ phoneme-labeled speech utterances which are partitioned into two sets: \textcolor{blue}{a train set consists of $4620$ utterance, and a test set consists of $1680$ utterances.} There are totally $61$ different types of phones in TIMIT labels. Our experiments in this section are carried out in two different cases: i) full $61$-phoneme classification case, where both training and testing use the $61$-phoneme data; ii) $39$-phonme classification case, where $61$ phonemes are folded into $39$ phonemes as described in \cite{Perdigao11}. \textcolor{blue}{The mapping table for the folding process is attached in appendix.}

@@

\begin{table}
  \caption{Test accuracy table for $39$ dimensional features }\label{tab:acc-classification39f}
  \vspace{-0.3cm}
  \centering
  \subcaption{Folded $39$ Phonemes}\label{tab:acc-classification39f_a}
  \centering  
  \begin{tabular}{l|c|c|c}
    \toprule
    {Model} &  K=1 &  K=3 &  K=5  \\  \midrule
    GMM-HMM & $62.3\%$ &  $68.0\%$ &  $68.7\%$  \\ \midrule
    GenHMM & $76.7\%$ & $77.7\%$ &  $77.7\%$ \\ 
    \bottomrule                                                                  
  \end{tabular}
  \bigskip
  \subcaption{$61$ Phonemes}\label{tab:acc-classification39f_b}
  % \centering  
  \begin{tabular}{l|c|c|c} \toprule
    {Model} &  K=1 &  K=3 &  K=5
    \\ \midrule
    GMM-HMM & $53.6\%$ &  $59.6\%$ & $61.9\%$  \\ \midrule
    GenHMM & $69.5\%$ & $70.6\%$ & $70.7\%$   \\ 
    \bottomrule
  \end{tabular}
  \vspace{0.3cm}
\end{table}

We firstly show the phoneme classification using 39 dimensional MFCC features (MFCC coefficients, deltas, and delta-deltas), to validate one possible usage of our proposed model. Since generative training is carried out in our experiments, GMM-HMM is trained and tested as a reference model in our experiments. Training and testing of GMM-HMM is in the same condition as GenHMMs are trained and tested. Dataset usage for GenMM and GMM-HMM is the same, and number of states for GMM-HMM is the same as that for GenHMM in modeling each phoneme. Apart from setting the reference model, we also run the experiment comparisons with different total number of mixture components ($K$).

@@

\begin{table}
  \caption{Test accuracy table for $13$ dimensional features }
  \label{tab:acc-classification13f}
  \vspace{-0.3cm}
  \centering
  \subcaption{Folded $39$ Phonemes}
  \centering  
  \begin{tabular}{l|c|c|c}
    \toprule
    {Model} &  K=1 &  K=3 &  K=5  \\  \midrule
    GMM-HMM & $48.5\%$ &  $51.2\%$ &  $52.4\%$  \\ \midrule
    GenHMM & $61.1\%$ & $62.1\%$ & $62.1\%$   \\ 
    \bottomrule                                                                  
  \end{tabular}
  \bigskip
  \subcaption{$61$ Phonemes}
  \begin{tabular}{l|c|c|c}
    \toprule
    {Model} &  K=1 &  K=3 &  K=5
    \\ \midrule
    GMM-HMM & $37.1\%$ &  $40.6\%$ & $42.2\%$  \\ \midrule
    GenHMM & $50.3\%$ & $51.0\%$ & $52.3\%$   \\ 
    \bottomrule
  \end{tabular}
  \vspace{0.3cm}
\end{table}

@@

>>>>

The problem \eqref{eq:em-m-opt} can be reformulated as
\begin{align}\label{eq:m-step-subs}
  &\umax{\bm{H}} \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) \nonumber \\
  =&\umax{\bm{q}}\Qq(\bm{q}; \bm{H}^{\mathrm{old}}) + \umax{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}}) 
     + \umax{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}),
\end{align}
where
\begin{align}
  \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) 
  % &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p({s}_{1})  \right] \nonumber\\
    &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p({s}_{1};\bm{H})  \right], \label{eq:init-distribution-update}\\
  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})}\hspace{-0.1cm}\left[ \sum_{t=1}^{T}\log\,p({s}_{t+1}|{s}_{t}; \bm{H}) \right], \label{eq:transition-update}\\
  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p(\ubm{x}, \ubm{\kappa}| \ubm{s}; \bm{H}) \right]. \label{eq:generative-model-update}
\end{align}

@@

For a flow model, let us assume that
the feature $\bm{h}_l$ at the $l$'th layer has two subparts as
$\bm{h}_l = [\bm{h}_{l,a}^{\intercal} \, , \, \bm{h}_{l,b}^{\intercal}]^{\intercal}$. The efficient invertible mapping of flow model comes from following forward and inverse relations between $(l-1)$'th and $l$'th layers
\begin{align}\label{eq-gl}
    \bm{h}_{l} &=
    \begin{bmatrix}
      \bm{h}_{l,a}\\
      \bm{h}_{l,b}
    \end{bmatrix}
    =
    \begin{bmatrix}
      \bm{h}_{l-1,a}\\
      \left(  \bm{h}_{l-1,b} - \bm{m}_b(\bm{h}_{l-1,a}) \right)\oslash \bm{m}_a(\bm{h}_{l-1,a}) 
    \end{bmatrix}, \nonumber\\
        \bm{h}_{l-1} \hspace{-0.1cm}&=\hspace{-0.1cm}
    \begin{bmatrix}
      \bm{h}_{l-1,a}\\
      \bm{h}_{l-1,b}
    \end{bmatrix}
    \hspace{-0.1cm}=\hspace{-0.1cm}
    \begin{bmatrix}
      \bm{h}_{l,a}\\
      \bm{m}_a(\bm{h}_{l,a})\odot \bm{h}_{l,b} + \bm{m}_b(\bm{h}_{l,a})
    \end{bmatrix},
\end{align}
where $\odot$ denotes element-wise product, $\oslash$ denotes
element-wise division, and $\bm{m}_a(\cdot), \bm{m}_b(\cdot)$ can be
complex non-linear mappings (implemented by neural networks).
For the flow model, the determinant of Jacobian matrix is
\begin{equation}\label{eq:cat-jacobian}
  \begin{array}{rl}
    \mathrm{det}(\nabla{\bm{f}_{s,\kappa}}) & = \prod_{l=1}^L \det (\nabla{\bm{f}_{s,\kappa}^{[l]}}),
  \end{array}
\end{equation}
where $\nabla{f_{s,\kappa}^{[l]}}$ is the Jacobian of the mapping from the $l$-th layer to the $(l-1)$-th layer, i.e., the inverse transformation. We compute the determinant of the Jacobian matrix as
\begin{align}\label{eq-hl-determinate}
  \det (\nabla{f_{s,\kappa}^{[l]}})& = \det \left[  \pd{\bm{h}_{l-1}}{\bm{h}_l} \right] \nonumber\\
                                   & = \det
                                     \begin{bmatrix}
                                       \bm{I}_a & \mathbf{0} \nonumber\\
                                       \pd{\bm{h}_{l-1,b}}{\bm{h}_{l,a}} & \mathrm{diag}(\bm{m}_a(\bm{h}_{l,a}))
                                     \end{bmatrix}\nonumber\\
                                   &= \det \left( \mathrm{diag}(\bm{m}_a(\bm{h}_{l,a})) \right),
\end{align}
where $\bm{I}_a$ is identity matrix and $\mathrm{diag}(\cdot)$ returns a square matrix with the elements of $(\cdot)$ on the main diagnal.
% \textcolor{red}{do I still need this?}
% Then the pdf is
% \begin{align}
%   p(\bm{x}) & =  p(\bm{z}) \big| \mathrm{det}(\bm{J}) |_{\bm{z}={\bm{f}}({\bm{x}})}\big| \nonumber\\
%             &  = p(\bm{z}) \prod_{l=1}^L \abs{\det\left( \mathrm{diag}(\bm{m}_a(\bm{h}_{l,a}))]  \right)}.
                %   \end{align}

@@

$p(\smtr{s}{1}{r};\bm{H})$ is the probability of initial state of GenHMM for $r$-th sequential sample. Actually $q_i = p({s}_{1} =i;\bm{H}) $, $i= 1, 2, \cdots, |\Ss|$. Solution to the problem
\begin{align}
  \bm{q} \hspace{-0.1cm} = \hspace{-0.1cm} \uargmax{\bm{q}} \Qq(\bm{q}; \bm{H}^{\mathrm{old}}),\; \mathrm{s.t.} \sum_{i=1}^{ |\Ss| }q_i = 1, q_i \geq 0, \forall i.
\end{align}
is
\begin{equation}\label{eq:update-initial-state-prob}
  q_i = \frac{1}{R} \sum_{r=1}^{R} p(\smtr{s}{1}{r}=i | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}), \forall\; i = 1, 2, \cdots, |\Ss|.
\end{equation}

@@

Since $\bm{A}_{i, j}  = p(\smtr{s}{t+1}{r}=j|\smtr{s}{t}{r}=i; \bm{H})$ is the element of transition matrix $\bm{A}$, the solution to the problem
\begin{align}\label{eq:update-transition-prob}
  \bm{A} = &\uargmax{\bm{A}} \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) \nonumber \\
  \mathrm{s.t.} &\hspace{0.2cm} \bm{A} \cdot \bm{1} = \bm{1}, \bm{A}_{i,j} \geq 0 \,\, \forall i,j,
\end{align}
is
\begin{equation}
  \bm{A}_{i,j} = \frac{\bar{\xi}_{i,j}}{\sum_{k = 1}^{|\Ss|} \bar{\xi}_{i,k}},
\end{equation}
where
\begin{equation}\label{eq:update-transition-solt}
  \bar{\xi}_{i,j} = \sum_{r= 1}^{R} \sum_{t= 1}^{T^{r}-1}{p(\smtr{s}{t}{r}=i, \smtr{s}{t+1}{r}=j| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}.
\end{equation}
\subsection{On Convergence of GenHMM}
In pursuit of representing a dataset by GenHMM,  we are interested if the learning solution discussed in subsection~\ref{subsec:optmGenHMM} would convergence. The properties on GenHMM's convergence is analyzed as follows.

@@

\subsection{Algorithm of GenHMM}
\begin{algorithm}[H]
  \caption{Learning of GenHMM}\label{algo:genhmm}
  \begin{algorithmic}[1]
    \STATE {\bfseries Input:}{
      Initializing $\bm{H}^{\mathrm{old}}, \bm{H} \in \Hh$ gives: \\
      $\bm{H}^{\mathrm{old}} = \{\Ss, \bm{q}^{\mathrm{old}}, A^{\mathrm{old}}, p(\bm{x}|s; \bm{\Phi}^{\mathrm{old}})\}$, $\bm{H} = \{\Ss, \bm{q}, A, p(\bm{x}|s; \bm{\Phi})\}$,\\
      empirical distribution $\hat{p}(\bm{x})$ of dataset;} \\
    \STATE Initialize $\bm{H}$
    \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
    \STATE Set learning rate $\eta$, neural network optimization epochs $N$ per EM step
    \FOR { $\bm{H}$ not converge}
    \FOR {epoch $n < N$}
    \STATE Sample a batch of data $\left\{ \ubmr{x}{r} \right\}_{r=1}^{R_b}$ from dataset $\hat{p}(\ubm{x})$ with batch size $R_b$

@@

@@

The TIMIT dataset consists of $5300$ phoneme-labeled speech utterances which are partitioned into two sets: {a train set consists of $4620$ utterance, and a test set consists of $1680$ utterances.} There are totally $61$ different types of phones in TIMIT labels. Our experiments in this section are carried out in two different cases: i) full $61$-phoneme classification case, where both training and testing use the $61$-phoneme data; ii) $39$-phonme classification case, where $61$ phonemes are folded into $39$ phonemes as described in \cite{Perdigao11}.

@@

\begin{table}
  \caption{Test accuracy table for $39$ dimensional features }\label{tab:acc-classification39f}
  \vspace{-0.3cm}
  \centering
  \subcaption{Folded $39$ Phonemes}\label{tab:acc-classification39f_a}
  \vspace{-0.2cm}
  \centering  
  \begin{tabular}{l|c|c|c}
    \toprule
    {Model} &  K=1 &  K=3 &  K=5  \\  \midrule
    GMM-HMM & $62.3\%$ &  $68.0\%$ &  $68.7\%$  \\ \midrule
    GenHMM & $76.7\%$ & $77.7\%$ &  $77.7\%$ \\ 
    \bottomrule                                                                  
  \end{tabular}
  \subcaption{$61$ Phonemes}\label{tab:acc-classification39f_b}
  \vspace{-0.2cm}
  % \centering  
  \begin{tabular}{l|c|c|c} \toprule
    {Model} &  K=1 &  K=3 &  K=5
    \\ \midrule
    GMM-HMM & $53.6\%$ &  $59.6\%$ & $61.9\%$  \\ \midrule
    GenHMM & $69.5\%$ & $70.6\%$ & $70.7\%$   \\ 
    \bottomrule
  \end{tabular}
  \vspace{0.2cm}
\end{table}

We firstly show the phoneme classification using 39 dimensional MFCC features (MFCC coefficients, deltas, and delta-deltas), to validate one possible usage of our proposed model. Since generative training is carried out in our experiments, GMM-HMM is trained and tested as a reference model in our experiments. Training and testing of GMM-HMM is in the same condition as GenHMMs are trained and tested. Dataset usage for GenMM and GMM-HMM is the same, and number of states for GMM-HMM is the same as that for GenHMM in modeling each phoneme. Apart from setting the reference model, we also run the experiment comparisons with different total number of mixture components.

@@

\begin{table}
  \caption{Test accuracy table for $13$ dimensional features }
  \label{tab:acc-classification13f}
  \vspace{-0.3cm}
  \centering
  \subcaption{Folded $39$ Phonemes}
  \vspace{-0.2cm}
  \centering  
  \begin{tabular}{l|c|c|c}
    \toprule
    {Model} &  K=1 &  K=3 &  K=5  \\  \midrule
    GMM-HMM & $48.5\%$ &  $51.2\%$ &  $52.4\%$  \\ \midrule
    GenHMM & $61.1\%$ & $62.1\%$ & $62.1\%$   \\ 
    \bottomrule                                                                  
  \end{tabular}
  \subcaption{$61$ Phonemes}
  \vspace{-0.2cm}
  \begin{tabular}{l|c|c|c}
    \toprule
    {Model} &  K=1 &  K=3 &  K=5
    \\ \midrule
    GMM-HMM & $37.1\%$ &  $40.6\%$ & $42.2\%$  \\ \midrule
    GenHMM & $50.3\%$ & $51.0\%$ & $52.3\%$   \\ 
    \bottomrule
  \end{tabular}
  \vspace{0.2cm}
\end{table}

@@

\begin{table}
  \caption{Test accuracy table of perturbation with white noise ($K=3$, folded $39$ phonomes).}
  \label{tab:acc-classification39f_noise_snr}
  \vspace{-0.3cm}
  \centering
  \begin{tabular}{l|c|c|c|c}
    \toprule
    {SNR} &  15dB &  20dB &  25dB & 30 dB  \\  \midrule
    GMM-HMM & $36.6\%$ &  $44.2\%$ &  $50.8\%$ & $57.1\%$  \\ \midrule
    GenHMM & $52.4\%$ & $62.0\%$ &  $69.7\%$ & $74.3\%$ \\ 
    \bottomrule                                                                  
  \end{tabular}
  \vspace{0.2cm}
\end{table}
\begin{table}
  \caption{Test accuracy table of perturbation by different type of noise (SNR=$20$dB, $K=3$, folded $39$ phonomes).}
  \label{tab:acc-classification39f_noise_type}
  \vspace{-0.3cm}
  \centering
  \begin{tabular}{l|c|c|c|c}
    \toprule
    {Noise Type} &  White &  Pink &  Babble & Volvo  \\  \midrule
    GMM-HMM & $44.2\%$ &  $48.8\%$ &  $57.7\%$ & $66.6\%$  \\ \midrule
    GenHMM & $62.0\%$ & $65.1\%$ &  $70.0\%$ & $75.7\%$ \\ 
    \bottomrule                                                                  
  \end{tabular}
  \vspace{0.2cm}
\end{table}

Apart from standard classification testing, we also test the robustness of our model to noise perturbations. We train GenHMM with $K=3$ by clean TIMIT training data in the case of folded $39$ phonemes with $39$ dimensional features. The testing dataset is perturbed by either the same type of noise with different signal-to-noise ratio (SNR) as shown in table~\ref{tab:acc-classification39f_noise_snr}, or different type of noises with the same SNR as shown in table~\ref{tab:acc-classification39f_noise_type}. The noise data is from NoiseX database. The baseline of these two sets of experiments is the accuracy testing of GenHMM and GMM-HMM on clean testing data in the same experimental condition, where GenHMM has $77.7\%$ and GMM-HMM gets $68.8\%$ as shown in table~\ref{tab:acc-classification39f}. It is shown in table~\ref{tab:acc-classification39f_noise_snr} that GMM-HMM's performance degenerates more than GenHMM's performance at them same level of noise perturbation, though the accuracy of both models increases along the increase of SNR. Especially, for SNR=$30$dB, the accuracy of GenHMM drops only about $3\%$ (from $77.7\%$ to $74.3\%$), while GMM-HMM encounters more than $10\%$ decrease (from $68.0\%$ to $57.1\%$) due to the noise perturbation. In table~\ref{tab:acc-classification39f_noise_type}, the SNR remains constant and GenHMM is tested with perturbation of different noise types. It is shown that GenHMM still has higher accuracy at different types of noise perturbations than GMM-HMM. Among these four types of noise, white noise shows most significant impact to GenHMM while the impact of volvo noise is negligible.
